### 2022-02-24 paper.Rmd
extending kplus logic to higher order moments and covariances

---
title             : "k-plus: a bicriterion extension of k-means anticlustering"
shorttitle        : "k-plus anticlustering"

@@

    Anticlustering refers to a collection of methods that partition data sets into disjunct groups while aiming for two objectives: (a) high between-group similarity and (b) high within-group heterogeneity. For some anticlustering formulations, these two objectives coincide, i.e., high between-group similarity is achieved when within-group heterogeneity is maximized. The present paper focuces on maximizing between-group similarity by introducing the k-plus objective, a novel bicriterion for anticlustering. K-plus anticlustering is an extension of classical k-means anticlustering. While k-means anticlustering focuses on minimizing between-group-differences with regard to the means of attribute values, k-plus anticlustering includes a term to optimize similarity with regard to the variance at the same time. The k-plus bicriterion can be reduced to an augmentation of the input data: For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the original variable. Thus, k-plus leaves the original k-means criterion unchanged and only differs in the input data. A computer simulation and practical examples on real norming data show that k-plus anticlustering achieves high similarity both with regard to mean attribute values as well as the variance, usually with no noticable sacrifice to either criterion. The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity. K-plus anticlustering is easily accessible as it has been implemented as part of the open source R package `anticlust` and is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).

@@

keywords          : "Anticlustering, k-means, k-plus, variance, bicriterion optimization"
wordcount         : "X"

@@

To optimize overall group similarity---and hence to overcome the limitations of k-means anticlustering---I present k-plus anticlustering in this paper. The k-plus criterion is an extension of the k-means objective that includes a term that quantifies similarity with regard to the variance of data across groups. Interestingly, this bicriterion reformulation can be understood as an augmentation of the original data matrix, leaving the original k-means objective unchanged. I show that the k-plus criterion outperforms other anticlustering criteria when the goal is to minimize differences with regard to the mean and the variance of groups. In most cases, both objectives can be optimized satisfactorily at the same time without any loss on either. Moreover, in two pratical examples I illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].

@@

>>>>

---
title             : "k-plus Anticlustering: An Improved k-means Criterion for Maximizing Between-Group Similarity"
shorttitle        : "k-plus anticlustering"

@@

    Anticlustering refers to a collection of methods that partition data sets into disjoint groups while aiming for high between-group similarity and high within-group heterogeneity. Anticlustering is usually approached by optimizing a reversed clustering objective such as the k-means criterion. Introducing k-plus, this paper presents an extension of k-means to maximize between-group similarity. While k-means anticlustering minimizes between-group-differences only with regard to mean attribute values, k-plus anticlustering can be used to minimize between-group-differences with regard to the variance, covariance structure, and higher order moments. To minimize differences in variance, each variable is duplicated by computing the squared deviation between the original values and the variable's mean. Because k-means anticlustering minimizes differences with regard to variables' means, appending the duplicated variables to the original data set---and then applying standard k-means anticlustering---suffices to equalize the variance between groups. Following a similar logic, differences in skewness, kurtosis and covariances can be minimized. Thus, while k-plus is a separate criterion for anticlustering, it can be reduced to an augmentation of the input data while the original k-means objective remains unchanged. A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity with regard to mean attribute values as well as the variance and higher order moments, usually with no noticable sacrifice to either criterion. The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity. Example are given on how k-plus anticlustering can be applied using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).

@@

keywords          : "Anticlustering, k-means, k-plus, variance, covariance, skewness, kurtosis, higher order moments"
wordcount         : "X"

@@

To optimize overall group similarity---and hence to overcome the limitations of k-means anticlustering---I present k-plus anticlustering in this paper. The k-plus criterion is an extension of the k-means objective that includes a term that quantifies similarity with regard to the variance, covariances, and higher order moments of data across groups. Interestingly, this extended criterion can be understood as an augmentation of the original data matrix, leaving the original k-means objective unchanged. I show that the k-plus criterion outperforms other anticlustering criteria when the goal is to minimize differences with regard to the mean and the variance of groups. In most cases, both objectives can be optimized satisfactorily at the same time without any loss on either. Moreover, in two pratical examples I illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust`, an extension to the popular statistical programming language R [@papenberg2020; @R-base; @R-anticlust].

@@

## Skewness and kurtosis

## Covariance structure
